On Sample Size for Block Bootstrap Confidence Intervals to Have Desired Coverage Rates

### 2023-05-30 Manuscript/manuscript.tex
5/30 Manuscript Update

The necessary sample size for bootstrap standard errors to provide valid
uncertainty measures in practice has not been extensively studied. For
independent data and non-block bootstrap, \citet{hesterberg2015teachers} notes
that while percentile-based CIs from nonparametric bootstrap
are more accurate than $t$-intervals for larger sample sizes, they are
less accurate for smaller sample sizes. \citet{chernick2009revisiting} finds that the the best choice of method for estimating
the parameter of a distribution is dependent on the sample size, the number of
bootstrap resamples, and the confidence level. In the context of structural equation
modeling, \citet{nevitt2001performance} find
that a sample size of 200--1000 is usually sufficient for interval estimation
using standard nonparametric bootstrap. For dependent data, the necessary
sample
size for bootstrap is even less clear. In the setting of linear regression and focused on estimating the proportion of 
variation attributable to a certain variable, \citet{burch2012nonparametric} finds
that three nonparametric bootstrap methods (Standard, Percentile, and Bias-Corrected and
Accelerated) approach the coverage of a pivotal quantity as $n$ gets larger for an
underlying normal distribution. However, they find that for other distributions, 
coverage may deteriorate as $n$ increases. 
When attempting to create interval estimations for the Pearson correlation 
coefficient of bivariate normal data, \citet{puth2015variety} find bootstrap 
methods to be less accurate than the Fisher's transformation method even 
for a sample of size $n = 100$ and correlation coefficient equivalent to 0.
Few studies have provided a practical guide on what size is needed
for block bootstrap inference for dependent data.
In a linear regression setting
with
dependent data where the
regression errors are generated from a homoscedastic AR(1) process, 
\citet{goncalves2005bootstrap} find that standard error
estimates from moving 
block bootstrap in small samples may be more accurate than
inference from closed-form asymptotic estimates, but moving block bootstrap 
\jy{did they do block or standard bootstrap?}
\mc{moving block bootstrap}
CIs still do not have sufficient coverage of the target parameter even for sample size
$n = 1024$.
\jy{I added more references to the reference folder that studied coverage of
  bootstrap CIs but not block bootstrap CIs.}

@@

We designed a simulation study to compare the performance of the different
block
bootstrap CI methods. In particular, we generated time series $X_t$
from a 1st order auto-regressive (AR1) process:
\[
X_t = \phi X_{t-1} + \epsilon_t,
\]
where $\phi$ is an auto-regressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: $\{-0.4, -0.2, 0.0, 0.2, 0.4\}$.
The series $X_t$ has mean zero and variance
$\sigma_x^2 = (1 - \phi^2) \sigma_{\epsilon}^2$, so for each value of $\phi$,
we
set $\sigma_{\epsilon}^2 = 1 / (1 - \phi^2)$ such that $\sigma_x^2 = 1$.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/pct}
  \caption{}
  \label{fig:pct}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/prop}
  \caption{}
  \label{fig:prop}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/bc}
  \caption{}
  \label{fig:bc}
\end{figure}

@@

Again, this analysis \eds{what?} \mc{addressed} relies on the assumption that
there is a size $n$ large enough for the method to work: that is, the method's
performance improves as $n$ increases. Out of the five types of intervals used
in this study, this assumption was found to hold true with respect to
estimating $\phi$ only for Standard, Student's $t$, and Proposed Method CIs, whereas 
Percentile, CBP, $BC$, and $BC_a$ intervals exhibited coverage deterioration as $n$
increased. The Percentile CI's coverage deterioration can be attributed to
bias that is not corrected as $n$ increases. Specifically, as $n$ increases, 
the width of the CI decreases, but because the Percentile CI
underestimates $\phi$, the coverage decreases. The $BC_a$ CI does seem to correct the bias, but the acceleration factor seems to fail, as the width of the CI
seems to be inconsistent.

>>>>

The necessary sample size for bootstrap standard errors to provide valid
uncertainty measures in practice has not been extensively studied. For
independent data and non-block bootstrap, \citet{hesterberg2015teachers} notes
that while percentile-based CIs from nonparametric bootstrap
are more accurate than $t$-intervals for larger sample sizes, they are
less accurate for smaller sample sizes. \citet{chernick2009revisiting} finds that the the best choice of method for estimating
the parameter of a distribution is dependent on the sample size, the number of
bootstrap resamples, and the confidence level. In the context of structural equation
modeling, \citet{nevitt2001performance} find
that a sample size of 200--1000 is usually sufficient for interval estimation
using standard nonparametric bootstrap. For dependent data, the necessary
sample
size for bootstrap is even less clear. In the setting of linear regression and focused on estimating the proportion of 
variation attributable to a certain variable, \citet{burch2012nonparametric} finds
that three nonparametric bootstrap methods (Standard, Percentile, and Bias-Corrected and
Accelerated) approach the coverage of a pivotal quantity as $n$ gets larger for an
underlying normal distribution. However, they find that for other distributions, 
coverage may deteriorate as $n$ increases. 
When attempting to create interval estimations for the Pearson correlation 
coefficient of bivariate normal data, \citet{puth2015variety} find bootstrap 
methods to be less accurate than the Fisher's transformation method even 
for a sample of size $n = 100$ and correlation coefficient equivalent to $0$.
Few studies have provided a practical guide on what size is needed
for block bootstrap inference for dependent data.
In a linear regression setting
with
dependent data where the
regression errors are generated from a homoscedastic AR(1) process, 
\citet{goncalves2005bootstrap} find that standard error
estimates from moving 
block bootstrap in small samples may be more accurate than
inference from closed-form asymptotic estimates, but moving block bootstrap 
\jy{did they do block or standard bootstrap?}
\mc{moving block bootstrap}
CIs still do not have sufficient coverage of the target parameter even for sample size
$n = 1024$.
\jy{I added more references to the reference folder that studied coverage of
  bootstrap CIs but not block bootstrap CIs.}

@@

We designed a simulation study to compare the performance of the different
block
bootstrap CI methods. In particular, we generated time series $X_t$
from a 1st order auto-regressive (AR(1)) process:
\[
X_t = \phi X_{t-1} + \epsilon_t,
\]
where $\phi$ is an auto-regressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: $\{-0.4, -0.2, 0.0, 0.2, 0.4\}$.
The series $X_t$ has mean zero and variance
$\sigma_x^2 = (1 - \phi^2) \sigma_{\epsilon}^2$, so for each value of $\phi$,
we
set $\sigma_{\epsilon}^2 = 1 / (1 - \phi^2)$ such that $\sigma_x^2 = 1$.

@@

These results beg the question of why certain methods perform poorly. To address this question, plots of 100 CIs using each of the Percentile, $BC$, and Proposed methods were constructed.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/pct}
  \caption{100 replicate Percentile Bootstrap CIs for samples of $n = 800$ for
  an AR(1) process with $\phi = 0.4$. For each replicate, the lower and upper bounds of the CIs are displayed, as well as $\hat\theta_n$ and $\bar\theta_n^{(b)}$. }
  \label{fig:pct}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/bc}
  \caption{100 replicate Bias-Corrected Bootstrap CIs for samples of $n = 800$ for
  an AR(1) process with $\phi = 0.4$. For each replicate, the lower and upper bounds of the CIs are displayed, as well as $\hat\theta_n$ and $\bar\theta_n^{(b)}$.}
  \label{fig:bc}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/prop}
  \caption{100 replicate Proposed Method CIs for samples of $n = 800$ for
  an AR(1) process with $\phi = 0.4$. For each replicate, the lower and upper bounds of the CIs are displayed, as well as $\hat\theta_n$ and $\bar\theta_n^{(b)}$.}
  \label{fig:prop}
\end{figure}

As shown in Figure~\ref{fig:phi}, Percentile CIs appear to suffer from bias; specifically, they seem to heavily underestimate $\phi$. As shown in Figure~\ref{fig:bc, $BC$ CIs do seem to correct this bias on average, but the width of the 
CIs is very inconsistent. From Figure~\ref{fig:prop, we can see that 
$prop$ CIs correct the bias using $\hat{bias} = \bar\theta_n^{(b)} - \hat\theta_n$, which seems to usually be negative, indicating a positive correction from
interval construction in \ref{sec:bbci}.

@@

Again, this analysis \eds{what?} \mc{addressed} relies on the assumption that
there is a size $n$ large enough for the method to work: that is, the method's
performance improves as $n$ increases. Out of the five types of intervals used
in this study, this assumption was found to hold true with respect to
estimating $\phi$ only for Standard, Student's $t$, and Proposed Method CIs, whereas 
Percentile, CBP, $BC$, and $BC_a$ intervals exhibited coverage deterioration as $n$
increased. The Percentile CI's coverage deterioration can be attributed to
bias that is not corrected as $n$ increases. Specifically, as $n$ increases, 
the width of the CI decreases, but because the Percentile CI
underestimates $\phi$, the coverage decreases. The $BC$ CI seems to correct the bias, but the width of the CI seems to be inconsistent. The acceleration 
factor of the $BC_a$ CI seems to fail, as the width of the CI
seems to be inconsistent.
